---
title: 'Exercise 2: ORGB 672'
author: "Lakshya Agarwal"
date: "`r Sys.Date()`"
output:
  pdf_document: 
    toc: true
    number_sections: true
    highlight: tango
  github_document:
    toc: true
---

\newpage

# Setup

```{r setup, echo=T, results='hide'}
library(tidygraph)
library(tidyverse)
library(igraph)
library(ggplot2)
library(vroom)
library(scales)
library(ggraph)
library(ggtext)
library(ggrepel)
library(ggforce)
library(ggthemes)
library(patchwork)
library(qualpalr)

# setwd("./Exercise 2")
```

# Code

## Create a graph with the specified edges  
```{r}
edges <- vroom("./edge_list.csv", delim = ",")

graph <- as_tbl_graph(edges, directed = FALSE)

graph
```


## Calculate centrality measures for {A, B, C, D}
```{r}
centrality_measures <- graph %>%
  activate(nodes) %>%
  mutate(
    degree = centrality_degree(),
    closeness = centrality_closeness(),
    betweenness = centrality_betweenness(),
    label = paste(
      "Name: ", name,
      "\nDegree: ", round(degree, 2),
      "\nCloseness: ", round(closeness, 2),
      "\nBetweenness: ", round(betweenness, 2)
    ),
    color = case_when(
      name %in% c("A", "B", "C", "D") ~ "red",
      TRUE ~ "black"
    )
  )

centrality_measures
```

## Plot network with labels
```{r fig.width=16, fig.height=9}
network_plot <- centrality_measures %>%
  ggraph(layout = "stress") +
  geom_edge_link(width = 1) +
  geom_node_point(
    size = 5,
    aes(color = color)
  ) +
  scale_color_manual(values = c("black", "red")) +
  geom_node_text(aes(label = label), size = 5, repel = TRUE) +
  theme_void() +
  theme(
    legend.position = "none",
    panel.border = element_rect(fill = NA)
  ) +
  plot_annotation(
    title = "Fakebook Bus Seat Selection",
    subtitle = "Possible seats: A, B, C, D",
    caption = "Data: Slides | Graphic: @lakshyaag"
  )

network_plot

ggsave("fakebook_bus_network_plot.png", network_plot,
  width = 16, height = 9, dpi = 300
)
```

# Fakebook Bus Seat Selection Discussion

As a new intern at Fakebook, choosing where to sit on the company bus is an important decision for fostering informal connections with coworkers.

## Seat A
- **Degree**: 3
- **Closeness**: 0.06
- **Betweenness**: 14

### Benefits
- **Central Role**: With the highest betweenness, consistently choosing Seat A will place the intern in a position where many paths pass through, making them a key connector in the network.
- **Influence**: The intern could potentially influence the flow of information and communication by taking a central seat.

### Drawbacks
- **Responsibility**: The central role might come with the pressure of being involved in many interactions.
- **Overload**: There's a risk of becoming overwhelmed by the need to facilitate conversations.

## Seat B
- **Degree**: 5
- **Closeness**: 0.07
- **Betweenness**: 12.08

### Benefits
- **Social Hub**: Seat B has the highest degree, offering the opportunity to directly connect with more people.
- **Accessibility**: High closeness means that the intern is, on average, closer to everyone else, facilitating easier communication and relationship-building.

### Drawbacks
- **Less Privacy**: More connections might mean less personal space and more interruptions.
- **High Engagement**: The intern might be expected to engage with a larger number of people regularly due to high proximity.

## Seat C
- **Degree**: 4
- **Closeness**: 0.07
- **Betweenness**: 6

### Benefits
- **Balanced Interaction**: Seat C offers a good balance of direct connections and centrality without the intensity of Seat B.
- **Strategic Position**: Moderately high closeness and betweenness suggest a strategic position for networking without being overwhelmed.

### Drawbacks
- **Less Prominent**: Not as central as Seat A or as connected as Seat B, which might limit networking opportunities.

## Seat D
- **Degree**: 4
- **Closeness**: 0.06
- **Betweenness**: 1.67

### Benefits
- **Quiet Networking**: Seat D allows for networking opportunities with a reasonable number of direct connections and a lower profile in the network.
- **Less Pressure**: Lower betweenness means less responsibility in facilitating interactions.

### Drawbacks
- **Peripheral Role**: Less centrality might make it harder for the intern to establish themselves as a key player in the network.
- **Indirect Influence**: The intern may have less influence on the overall flow of communication due to a lower betweenness centrality.

## Conclusion
In conclusion, the intern's choice should align with their networking style and desired level of interaction. 
If the goal is to be a central figure, Seat A or B might be the best choice. Seat C offers a balanced approach, while Seat D is not recommended for the intern's purposes.